Non-parametric approaches to optimal policy are more robust

### 2013-03-21 inst/doc/nonparametric-bayes.Rmd
writing writing

1. Classical approaches are to fix relationship of stock recruitment.  
2. Parameter uncertainty
3. Gaussian processes for model uncertainty

The problem of structural uncertainty in managing ecological systems.

(Big literature here of course, primary approaches have been using
parametric uncertainty and model uncertainty (e.g. belief distribution over models 
I believe. Worth distinguishing our problem from the more trivially 
sized problems in which the action space is discrete and consists
of 2 or three possible choices, e.g. much of the recent Possingham et al work.)

@@

Stochastic dynamic programming @Mangel1982

@@

Background on the concerns of structural uncertainty -- we don't have
the right models.  Further unknonws such as measurement uncertainty,
parameter uncertainty, unobserved states, knowledge of boundary
conditions, etc. further compound the issue.  Though a hierarchical
Bayesian approach provides a natural way to address these from a
statistical standpoint, formulating reasonable parametric descriptions
of each form of uncertainty is a challenging task in itself, let alone
the computational difficulties of solving such a system. @Cressie2009
provides a good account of the successess and challenges of the approach. 
Applying these approaches in the management context of sequential decision
making, in which forecasts must be obtained over a range of possible actions
and updated regularly as new information arrives makes such an approach 
less feasible still.

* Management goals / decision-theoretic approaches need accurate
prediction over relevant (short?) timescales more than accurate (but
incomplete or noisy estimated) mechanisms.

* Nonparametric (machine-learning?) approaches may offer the benefit of
the hierarchical Bayesian approach without the practical and computational
limitations of their parametric kin.  Non-parametric models are flexible
enough to take maximum advantage of the data available, while being
appropriately ambiguous about the dynamics of a system in regions of
parameter space that have been poorly or never sampled.

@@

.  However, such approaches have yet to be applied to the
decision-theoretic framework that could guide management decisions,
where we expect them to excel for several reasons (1) the ability to
make accurate forecasts by more closely approximating the underlying
process where data is available (2) remaining appropriately ambiguous
where data is not available (3) remaining computationally simple enough
to avoid some pitfalls common to hierarchical parametric approaches.

@@

The essense of the Gaussian process approach can be captured in the following thought experiment: An exhaustive parametric approach to the challenge of structural uncertainty might proceed by writing down all possible functional forms for the underlying dynamical system with all possible parameter values for each form, and then consider searching over this huge space to select the most likely model and parameters; or using a Bayesian approach, assign priors to each of these possible models and infer the posterior distribution of possible models. The Gaussian process approach can be thought of as a computationally efficient approximation to this approach. Gaussian processes represent a large class of models that can be though of as capturing or reasonably approximating the set of models in this collection.  By modeling at the level of the process, rather than the level of parametric equation, we can more concisely capture the possible behavior of these curves.  In place of a parametric model of the dynamical system, the Gaussian Process approach postulates a prior distribution of (n-dimensional) curves that can be though of as approximations to a range of possible (parametric) models that might describe the data. The GP allows us to consider a set of possible curves simultaneously (Figure 1).  <!-- Figure of curves drawn from the posterior density? -->

@@

* Non-parametric methods have received far too little attention in
ecological modeling efforts that are aimed at improved conservation
planning and decision making support.

* Importance of non-parametric approaches in conservation planning /
resource management / decision theory.

* Decision-theoretic tools such as optimal control calculations rely
on robust _forecasting_ more strongly than they rely on accurate
_mechanistic_ relationships.

@@

* Anticipate improved relative performance in higher dimensional examples

* Discuss constant escapement in model, in policies.

@@

* Discussion of alternative related approaches: POMDP/MOMDP,

Future directions
--------------------------------------------------------------------

>>>>

@@

<!-- Better defintion of focal problems here .... -->
Stochastic dynamic programming @Mangel1982

@@

Structural uncertainty.

Further unknowns such as measurement uncertainty, parameter uncertainty,
unobserved states, knowledge of boundary conditions, etc. further
compound the issue.  Though a hierarchical Bayesian approach provides a
natural way to address these from a statistical standpoint, formulating
reasonable parametric descriptions of each form of uncertainty is a
challenging task in itself, let alone the computational difficulties
of solving such a system. @Cressie2009 provides a good account of the
successess and challenges of the approach.  Applying these approaches in
the management context of sequential decision making, in which forecasts
must be obtained over a range of possible actions and updated regularly
as new information arrives makes such an approach less feasible still.

@@

<!-- Whoa, isn't machine learning used to make decisions all the time,
 eg, in engineering and Artificial Intelligence?  -->

A more immediate barrier to their adoption is the absence of a framework
for applying machine learning approaches to resource management problems.
Traditional approaches to optimal control (Pontryagin's principle, stochastic
dynamic programming) rely on knowledge of the state equation, usually described
by a simple parametric model. Here we illustrate how a stochastic dynamic
programming algorithm can alternatively be driven by the predictions from
a Gaussian process -- a machine learning approximation to the state dynamics.  
<!-- Okay, perhaps that's novel, but it's pretty trivial.  Isn't it
obvious to everyon that it's trivial? -->

<!-- paraphrase advantages of machine learning -->

Management goals / decision-theoretic approaches need accurate prediction
over relevant (short?) timescales more than reasonable long-term
mechanisms.  Machine-learning approaches may offer the benefit of the
hierarchical Bayesian approach without the practical and computational
limitations of their parametric kin.  Non-parametric models are flexible
enough to take maximum advantage of the data available, while being
appropriately ambiguous about the dynamics of a system in regions of
parameter space that have been poorly or never sampled.

@@

The essense of the Gaussian process approach can be captured in the
following thought experiment: An exhaustive parametric approach to the
challenge of structural uncertainty might proceed by writing down all
possible functional forms for the underlying dynamical system with all
possible parameter values for each form, and then consider searching
over this huge space to select the most likely model and parameters;
or using a Bayesian approach, assign priors to each of these possible
models and infer the posterior distribution of possible models. The
Gaussian process approach can be thought of as a computationally
efficient approximation to this approach. Gaussian processes represent
a large class of models that can be though of as capturing or reasonably
approximating the set of models in this collection.  By modeling at the
level of the process, rather than the level of parametric equation,
we can more concisely capture the possible behavior of these curves.
In place of a parametric model of the dynamical system, the Gaussian
Process approach postulates a prior distribution of (n-dimensional)
curves that can be though of as approximations to a range of possible
(parametric) models that might describe the data. The GP allows us
to consider a set of possible curves simultaneously.  
<!-- Figure 1 include curves drawn from the posterior density?  -->

@@

Non-parametric methods have received far too little attention in
ecological modeling efforts that are aimed at improved conservation
planning and decision making support.

@@

Three elements contribute to the machine learning model performing as well
or better than the mechanisitic models in this context.

### 1. Relevant state space

The dynamics occur over a range of state-space in which the machine learning
model closely matches the predictions of y
implausible model performs as well or better than the more plausible
model.

This is the most obvious and immediate reason why the shortcomings
that appear to make the model implausible do not make it useless.
This effect is enhanced in our example because the management actions help
drive the system towards rather than away from this region of state-space.

### 2. Predictive accuracy

This application relies only on the predictive accuracy of the model,
not an interpretation of the parameters.

<!-- wow, run-on tangent? -->
Predictive accuracy is not the goal of all modeling, as ecologists
have been observing for as long as they made models (perhaps none
more memorably than @Levins1969).  Mechanistic modeling is at its
most powerful not when it is used to directly forecast future states
but when it provides an understanding of how to approach a problem.
SIR-type models from the epidemiological literature are a good example.
While the simplest SIR models have little predictive power over the
outbreak intensity or timing at a particular location, they provide a
powerful understanding of the spread of an infection in terms of a single,
biologically meaningful parameter: $R_0$, the basic reproductive number.
From the model, it becomes clear that management need not vaccinate every
member of the population to stop the spread, but rather it suffices to
vaccinate a sufficient fraction of the population to reduce $R_0$ below 1.

### 3. Time scale for new data

In the sequential decision making problem we considered, we are
presented with new data after each action.  The relevant timescale
for the prediction is thus not the long-term dynamics, which would be
wildly divergent, but the dynamics over this much shorter interval.
While ecologists may be hesitant to base continual management on a model
with obviously inaccurate long-term behavior, engineers tend to consider
the problem in frequency space and gravitate to the opposite position --
a good control model should prioritize high-frequency accuracy over low
frequency accuracy.  The differences in intuition may arise from the
timescales at which each profession can typically adjust their control
variables -- much faster for a control system of a chemical plant than
state policy for a natural resource.  Still, the lesson is clear: when
facing repeated management decisions over a short timescale, such as
setting annual harvests of a fishery, it may be more valuable to use a
machine learning algorithm that makes accurate year-ahead predictions
that capture some of the high-frequency fluctuations that appear only
as noise in a mechanistic model of the long-term population dynamics.

2. Better express ambiguity in regions where data is not available

One of the greatest strengths of mechanistic models is their greatest weakness as well.

## Future directions

<!-- Jargony and vague -->
While we have highlighted certain generic of this problem that allow the
nonparametric approach to excel -- short timescales between successive
observations and actions, the accuracy in the appropriate region of state
space, the ability to express uncertainty outside the observed data --
there are equally several aspects for which the nonparametric approach is
at a relative disadvantage.

<!-- Model complexity -->
In this simulated example, the underlying
dynamics are truly governed by a simple parametric model, allowing
the parametric approaches to be more accurate.  Similarly, because the
dynamics are  one-dimensional dynamics and lead to  stable nodes (rather
than other attractors such as limit-cycles resulting in oscillations),
the training data provides relatively limited information about the
dynamics.  For these reasons, we anticipate that in higher-dimensional
examples characteristic of ecosystem management problems that the machine
learning approach will prove even more valuable.

<!-- Data complexity. Perhaps too far out of scope... -->
The machine learning approach is also better suited to complex and
disparate data.  Incorporating various sources of information into
mechanistic models can be an immensely difficult due to the increased
complexity involved.  Only thanks to tandem advances in increasing
computational power and hierarchical statistical methodology have we
been able to tackle such intuitively important complexity (and the
potentially new available data that accompanies it) such as spatial
distribution, heterogeneities of space, time, and individuals, to shift
to ecosystem-based approaches from single-species based approaches.
Without the need to formulate mechanisms, many modern machine learning
algorithms can leverage potential information from all available sources
of data directly.  The algorithms can recognize unanticipated or subtle
patterns in large data sets that enable more accurate predictions than
mechanistic models that are formulated at a more macroscopic level.

<!-- 
* Discuss constant escapement in model, in policies.

@@

* Discussion of alternative related approaches: POMDP/MOMDP
-->

--------------------------------------------------------------------
